---
title: "A2 - SQL and R"
author: "Tora Mullings"
date: "2/12/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Required libraries and installations.
You will need a local installation of pgadmin4 as well.
```{r}
library(DBI)
library(config)
library(RPostgres)
```

# Credentials
Place the attached config.yml file in your working directory. It will pass on the database credentials to the connection.

```{r}
options(warn=-1)
dw <- config::get("datawarehouse")
```

### In pgadmin4, create a database named 'movies'. Connect to the database.
```{r}
con <- dbConnect(RPostgres::Postgres(),
                 dbname = dw$database, 
                 host = dw$server,
                 port = dw$port,
                 user = dw$uid,
                 password = dw$pwd
                 )
```

# Table construction

### Create the 'ratings' table. 7 of my friends rated these 6 movies on a scale of 1-5.

```{r}
query <- dbSendQuery(con, query <- 'DROP TABLE IF EXISTS ratings;')
```

```{r}
query <- dbSendQuery(con, query <- '
CREATE TABLE ratings (
	name CHAR ( 15 ),
	Eternals INTEGER NOT NULL,
	Encanto INTEGER NOT NULL,
	TheFallout INTEGER NOT NULL,
	Cherry INTEGER NOT NULL,
  LastLooks INTEGER NOT NULL,
	NoTimetoDie INTEGER NOT NULL 
);'
)
```

### Insert the data into the ratings table. 
```{r}
query <- dbSendQuery(con, query <- "INSERT INTO 
    ratings (name, Eternals, Encanto, TheFallout, Cherry, LastLooks, NoTimetoDie)
  VALUES
  ('Char', 1, 4, -1, -1, 3, 3),
  ('Yang', 2, 4, 4, -1, 3, 3),
  ('Ory', -1, 5, 3, 2, 3, 4),
	('Bloom', 3, 4, -1, 3, 3, -1),
	('Rin', -1, 5, 6, -1, 3, 3),
	('Tach', -1, 4, 3, -1, 4, 4),
  ('Poppy', 3, 5, 3, 2, -1, -1); "
)
```

### View the table. There are 6 movies and 7 friends.
```{r}
df <- dbReadTable(con, "ratings")
head(df)
```

# Handling missing values

### Some friends haven't seen these movies, so there is -1 in those places in the table. We will replace these values with the mean of each movie column.

```{r}
df1 <- df
df1[df1 == -1] <- NA

impute_missing <- function(data, columns) {
  for(i in columns) {
    imputed <- round(mean(data[ , i], na.rm = TRUE), digits=2)
    data[ , i][is.na(data[ , i])] <- imputed
  }
  return(data)
}

```

```{r}
df1 <- impute_missing(df1, c(2:7))
head(df1, n=7L)
```

# Conclusion
The set of data imputation techniques that could have been used in this instance is limited because the dataset's small size. If it were much larger, we might have been able to simply drop the rows that were missing ratings. Or, if there was a movie in particular with few ratings (meaning that very few of my friends saw it), I could have dropped that movie column. 

Another option for handling the missing data would have been to replace the -1 values with the middle value on the rating scale. The scale was 1-5, so in this case the middle is 3.




